---
title: "USA Emissions Data: Modeling Activity"
author: "Daniel Sussman"
format: gfm
---

We're going to be looking at data that collected as part of a study on environmental justice and harmful emissions.
An article in [Nature Communications](https://www.nature.com/articles/s41467-023-43492-9) details the results of the study.

The data for this study as well as (most of?) the code for the analysis can be found on [Github](https://github.com/yanellinunez/USA_emissions_code).
We created a simplified version of the dataset for us to explore.


```{r}
suppressPackageStartupMessages(library(tidyverse))
nox_df <- read_rds("USA_emissions_simple.rds")
nox_df
```




We want to explore the relationship between the race percentage variables and the variable `nox_energy_relat_chg`.
We can recompute the difference and relative change using these equations.

```{r}
nox_df |> 
  select(gisjoin, year, starts_with("nox_ener")) |> 
  filter(gisjoin == nth(gisjoin, 10)) |> 
  mutate(d = lead(nox_ener) - nox_ener, t = nox_energy_diff / nox_ener * 100)
```

The `nox_energy_relat_chg` variable, gives the relative change in NOx pollutants from the year to year + 10.

Let's plot the NOx energy relative change versus the `perc_white` variable.


```{r}
ggplot(nox_df)
```




```{r}
ggplot(nox_df, aes(x = perc_white, y = nox_energy_relat_chg)) + 
  geom_smooth() +
  scale_x_sqrt()
```

We'll create an initial training/testing split and then fit a model with `nox_energy_relat_chg` as our response variable and `perc_white` as our predictor variable.

```{r}
library(tidymodels)
split <- initial_split(nox_df, prop = .8)
training<- training(split)
```

### Evaluating model assumptions

Use `ggplot` and `broom::augment` to recreate the first two plots below.

```{r}
plot(mod1)
```
